---
title: 'Titanite solubility: Flanagan + LEPR experiments'
author: "John C. Ayers"
output:
  html_document:
    df_print: paged
    toc: yes
bibliography: "TitaniteReferences.bib"
---

```{r set-global, include=FALSE, echo=FALSE}
# Suppress all error messages
knitr::opts_chunk$set(error = FALSE)
```


```{r setup, include=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, car, leaps, knitr, skimr, ggExtra, gvlma, boot)

```

### Input data

```{r}
AllExpts <- read_csv("AllExptsV3.csv")
# Contains 32 of 39 DF experiments
skimr::skim(AllExpts)
```

### Calculate concentrations of cations

```{r, echo=FALSE}
Cations <- AllExpts %>%
  mutate(
    Si = SiO2/60.0848,
    Al = 2*Al2O3/101.961,
    Na = 2*Na2O/61.979,
    K = 2*K2O/94.194,
    Ca = CaO/56.0794,
    Fe = FeO/71.8464,
    Mg = MgO/40.3044,
    Ti = TiO2/79.8988,
    eCa = (CaO/56.0794 - Al2O3/101.961 + K2O/94.194 + Na2O/61.979),
    CationSum = (Si + Al + Na + K + Ca + Fe + Mg + Ti),
    lnTiO2 = log(TiO2),
    Al2O3m = Al2O3/101.961,
    Na2Om = Na2O/61.979,
    K2Om = K2O/94.194,
    CaOm = CaO/56.0794
     )
```

### Calculate cation fractions

```{r, echo=FALSE}

CationFractions <- Cations
f <- match("Si", names(CationFractions))
l <- match("Ti", names(CationFractions))
for (i in f:l) {
  CationFractions[(i)] = CationFractions[(i)]/CationFractions$CationSum
}

```

### Calculate melt composition parameters

Formula for calculating melt composition parameter M [@Ryerson1987]

$$M = \frac{(Na + K + 2*Ca)}{(Si * Al)}$$

```{r, echo=FALSE}
CationFractions <- CationFractions %>%
  mutate(
    M = (Na + K + 2*Ca)/(Si*Al),
    FM = (Na + K + 2*(Ca + Fe + Mg))/(Si*Al),
    C = (10*eCa)/(Si*Al),
     A_CNK = Al2O3m/(CaOm + Na2Om + K2Om),
    A_NK = Al2O3m/(Na2Om + K2Om),
    A_C = Al2O3m/CaOm,
    TK = TC+273.15,
    InvTK = 1/TK,
    PInvTK = (Pkb*1000 - 1)*InvTK
    )
# Filter out experiments with no titanite in run products
CationFractionsNoTitanite <- filter(CationFractions, Titanite == "N")
CationFractionsTitanite <- filter(CationFractions, Titanite == "Y")
```

### Correlation matrix
- "select" function doesn't work if MASS library is loaded

```{r, echo=FALSE}
CorrTest <- select(CationFractionsTitanite, TiO2, lnTiO2, Pkb, Ca, M:PInvTK)
cor(CorrTest)
```

-   Considering lnTiO2, InvTK correlates best. Melt compositions parameters have similar correlation coefficients. 
-   Pkb and PInvTK show poor correlations, so drop from model.

### Subsets regression

-   to find optimal set of parameters

```{r, echo=FALSE}
leaps <- regsubsets(lnTiO2 ~ TK + M + FM + C + A_C+ InvTK, data = CationFractionsTitanite, nbest = 2) 
plot(leaps, scale = "adjr2")

```

-   A_C should be in model.

### Scatterplot matrices

```{r, echo=FALSE, results='hide'}
scatterplotMatrix(~ lnTiO2 + Ca + InvTK + TK + A_C, data = CationFractionsTitanite, groups = CationFractionsTitanite$Source, smooth = FALSE)
dev.copy(png,"ScatterPlotMatrixAllExptslnTiO2.png")
dev.off()   

```

### Preliminary model fits

```{r, echo=FALSE}
RegModel <- CationFractionsTitanite
fit <- lm(lnTiO2 ~ A_C + InvTK + TK, data = RegModel)
summary(fit)

```

- Tried TK and InvTK below and got the same r^2 values, so use InvTK.

### Refined model fit

```{r, echo=FALSE, results='hide', fig.show='hide'}
fit <- lm(lnTiO2 ~ A_C + InvTK, data = RegModel)
out <- summary(fit)
plot(fit)
```

### Test of lm model regression assumptions

```{r, echo=FALSE}
gvmodel <- gvlma(fit)
summary(gvmodel)
```

Two assumptions violated [@pena2006global].

Try bootstrapping to avoid assumptions of normal regression. The methods below are described in section 12.6.2 of [@Kabacoff2015]

### Preliminary bootstrap model fit

```{r, echo=FALSE}
RegModel <- CationFractionsTitanite
bs <- function(formula, data, indices) {
  d <- data[indices,]
  bootfit <- lm(formula, data = d)
  return(coef(bootfit))
}
set.seed(4329)
results <- boot(data = RegModel, statistic = bs, R=1000, formula = lnTiO2 ~ A_C + InvTK)
print(results)
```

Coefficients are the same, but errors are higher.

```{r}
rsq <- function(formula, data, indices) {
  d <- data[indices,]
  bootfit <- lm(formula, data = d)
  return(summary(bootfit)$r.square)
}
set.seed(4329)
results <- boot(data = RegModel, statistic = rsq, R=1000, formula = lnTiO2 ~ A_C + InvTK)
print(results)  
plot(results) # plots distribution of r-squared values.
```

- Mean r^2 0.86, the same as in the simple lm fit.

Instead of iteratively removing outliers, try using a fitting method that is robust to the effect of outliers: MASS::rlm. See pg. 393 of [@Wickham2017]

```{r}
library(MASS)
MASSfit <- rlm(lnTiO2 ~ A_C + InvTK, data = RegModel)
MASSout <- summary(MASSfit)
MASSout
plot(MASSfit)
```

- Can't use gvlma to test assumptions of rlm model. Don't get an r^2 value, 
- The residual standard error on the lm model was 0.397 compared to 0.426 for the rlm model. The coefficient errors are similar.
- Try removing outliers from lm model and then compare to rlm without outlier removal.
- The outliers for lm and rlm are the same for the first step.

### Store coefficients in new dataframe

```{r, echo=FALSE, results='hide'}
Coeff <- unname(coefficients(fit))
Int <- Coeff[1]
A_C <- Coeff[2]
InvTK <- Coeff[3]
AdjRsq <- out$adj.r.squared
```


```{r, echo=FALSE, results='hide', fig.show='hide'}
DelExpts <- c("DAC", "SpDis_4", "SpG_28")
RegModel <- filter(RegModel, !(Expt %in%  DelExpts))
fit <- lm(lnTiO2 ~ A_C + InvTK, data = RegModel)
out <- summary(fit)
plot(fit)
Coeff <- unname(coefficients(fit))
Int[(2)] <- Coeff[1]
A_C[(2)] <- Coeff[2]
InvTK[(2)] <- Coeff[3]
AdjRsq[(2)] <- out$adj.r.squared
```

```{r, echo=FALSE, results='hide', fig.show='hide'}
DelExpts <- c("ASI200", "SpG_12", "RHY")
RegModel <- filter(RegModel, !(Expt %in%  DelExpts))
fit <- lm(lnTiO2 ~ A_C + InvTK, data = RegModel)
out <- summary(fit)
plot(fit)
Coeff <- unname(coefficients(fit))
Int[(3)] <- Coeff[1]
A_C[(3)] <- Coeff[2]
InvTK[(3)] <- Coeff[3]
AdjRsq[(3)] <- out$adj.r.squared
```

### Plot coefficient values for lm and rlm models

```{r, echo=FALSE}
# Add Massfit as reference line?
ModelCoeffs <- as.data.frame(cbind(Int, A_C, InvTK)) 
ModelCoeffs$Iter <- row.names(ModelCoeffs)
MassCoeff <- unname(coefficients(MASSfit))
# MassCoeff[(4)] <- MASSout$adj.r.squared
for (i in 1:3) {
  print(ggplot(ModelCoeffs, aes_(x=~Iter, y=as.name(names(ModelCoeffs)[i]))) +
    geom_point() + 
     geom_hline(yintercept = MassCoeff[i], color = "red") 
  )
}

```

- The rlm coefficients are closest to the lm results without deletion of outliers (iteration = 1). So rlm really doesn't do a good job of reducing the effect of outliers.

### Influence plot

```{r, echo=FALSE, results='hide'}
influencePlot(fit, main="Influence Plot", sub="Circle size is proportional to Cook's distance")

```

Observations depicted as large circles may have a disproportionate influence on the parameter estimates of the model. Points with Hat values above 0.2 have high leverage.

### Calculate model TiO2 concentrations and errors

```{r, echo=FALSE}
# Model output table provides standard errors; convert to standard deviation by multiplying by sqrt of n = # of samples
n <- nrow(RegModel)
A_Ccoef <- out$coefficients[2,1]
A_Ccoefsd <- out$coefficients[2,2]*sqrt(n)
InvTcoef <- out$coefficients[3,1]
InvTcoefsd <- out$coefficients[3,2]*sqrt(n)
Int <- out$coefficients[1,1]
RegModel <- RegModel %>%
  mutate(
   lnTiO2sat = (A_Ccoef*A_C + InvTcoef*InvTK + Int),
    lnTiO2satSD = sqrt(A_Ccoef^2*A_Ccoefsd^2 + InvTcoef^2*InvTcoefsd^2),
   lower = lnTiO2sat - lnTiO2satSD,
   upper = lnTiO2sat + lnTiO2satSD
  )
```

### Plot model lnTiO2 saturation concentration versus measured TiO2

```{r, echo=FALSE}
# Removed errors because they are wrong.
p <- ggplot(RegModel, aes(x = lnTiO2, y = lnTiO2sat, shape = Source, color = Source)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme(legend.position = "left") +
  ylab(~paste("ln TiO"[2],"sat")) +
  xlab(~paste("ln TiO"[2],"meas"))
# add marginal distributions
p1 <- ggMarginal(p, type = "histogram")
print(p1)
ggsave("AllExptsA_C_ModelFit.svg", width = 2000, height = 2*618, units = "px", plot = p1)

```

### lnTiO2-A_C plot

```{r, echo=FALSE}
RegModel %>%
ggplot(aes(x=A_C, y=lnTiO2, color = Source)) +
  geom_point() +
  geom_smooth(method = lm, se = TRUE) +
  theme_bw() +
  ylab(~paste("ln TiO"[2],"sat"))
  
```

### lnTiO2-InvTK plot

```{r, echo=FALSE}
ggplot(RegModel, aes(x=InvTK, y=lnTiO2, color=A_C)) +
  geom_point()
```

### Model output table

```{r, echo=FALSE}
detach("package:MASS", unload = TRUE)
lnTiO2AllExptsModelTable <- select(RegModel, Source, Expt, A_C, InvTK, lnTiO2, lnTiO2sat)
write.csv(lnTiO2AllExptsModelTable, file = "lnTiO2AllExptsM_ModelTable.csv")
```

### Model titanite saturation test

Use model equation to test whether LEPR expts without titanite or rutile are undersaturated.

```{r, echo=FALSE}
LEPR_Expts <- read_csv("LEPR_ExperimentsV2.csv")
LEPRLiqs <- read_csv("LEPR_LiquidAnhyV4.csv")
LEPR <- inner_join(LEPR_Expts, LEPRLiqs, by = "Index")
LEPRnoTitRut <- filter(LEPR, !(grepl("tit|rut", Phases))) %>%
  mutate(
    FeOt = FeO + (2*71.85*Fe2O3)/159.69,
    TK = T_C + 273.15,
    InvTK = 1/TK,
    Tot = SiO2 + TiO2 + Al2O3 + FeOt + MgO + CaO + Na2O + K2O) %>%
  filter(Tot > 90)
LEPRnoTitRut <- select(LEPRnoTitRut, Experiment.x, InvTK, SiO2, TiO2, Al2O3, FeOt, MgO, CaO, Na2O, K2O, Tot)   %>%
  mutate(
    SiO2 = 100*SiO2/Tot,
    TiO2 = 100*TiO2/Tot,
    Al2O3 = 100*Al2O3/Tot,
    FeOt = 100*FeOt/Tot,
    MgO = 100*MgO/Tot,
    CaO = 100*CaO/Tot,
    Na2O = 100*Na2O/Tot,
    K2O = 100*K2O/Tot,
    Tot = SiO2 + TiO2 + Al2O3 + FeOt + MgO + CaO + Na2O + K2O,
    lnTiO2 = log(TiO2),
    Al2O3m = Al2O3/101.961,
    CaOm = CaO/56.0794,
    A_C = Al2O3m/CaOm
      )
```

#### Compare model and observations

```{r, echo=FALSE}
NoTitaniteModel <- select(LEPRnoTitRut, Experiment.x, lnTiO2, InvTK, A_C)
NoTitaniteModel$lnTiO2sat <- predict(fit, NoTitaniteModel)
CompareTiO2 <- select(NoTitaniteModel, Experiment.x,  lnTiO2, lnTiO2sat)
CompareTiO2 <- CompareTiO2 %>%
  mutate(
    Model_Correct = if_else(lnTiO2sat > lnTiO2, "true", "false")
  )
CompareTiO2
CompareTiO2 %>%
  count(Model_Correct)
```

-   Model correctly predicts undersaturation in 91% of experiments.

### Save output file and display session info

```{r, echo=FALSE}
save.image("TitaniteAllExptslnTiO2A_C.RData")
sessionInfo()
```

### References
