---
title: 'Titanite solubility: Flanagan + LEPR experiments'
author: "John C. Ayers"
output:
  html_document:
    df_print: paged
    toc: yes
bibliography: "TitaniteReferences.bib"
---
```{r set-global, include=FALSE}
# Suppress all error messages
knitr::opts_chunk$set(error = FALSE)
```


```{r setup, echo=FALSE, include=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, car, leaps, knitr, skimr, ggExtra, gvlma)
AllExpts <- read_csv("AllExptsV2.csv")
# Contains 32 of 39 DF experiments
skimr::skim(AllExpts)
```

### Calculate concentrations of cations

```{r}
Cations <- AllExpts %>%
  mutate(
    Si = SiO2/60.0848,
    Al = 2*Al2O3/101.961,
    Na = 2*Na2O/61.979,
    K = 2*K2O/94.194,
    Ca = CaO/56.0794,
    Fe = FeO/71.8464,
    Mg = MgO/40.3044,
    Ti = TiO2/79.8988,
    CationSum = (Si + Al + Na + K + Ca + Fe + Mg + Ti)
     )
```

### Calculate cation fractions

```{r}

CationFractions <- Cations
f <- match("Si", names(CationFractions))
l <- match("Ti", names(CationFractions))
for (i in f:l) {
  CationFractions[(i)] = CationFractions[(i)]/CationFractions$CationSum
}

```

### Calculate melt composition parameters

Formula for calculating melt composition parameter M [@Ryerson1987]

$$M = \frac{(Na + K + 2*Ca)}{(Si * Al)}$$

```{r}
CationFractions <- CationFractions %>%
  mutate(
    M = (Na + K + 2*Ca)/(Si*Al),
    FM = (Na + K + 2*(Ca + Fe + Mg))/(Si*Al),
    TK = TC+273.15,
    InvTK = 1/TK,
    PInvTK = (Pkb*1000 - 1)*InvTK
    )
# Filter out experiments with no titanite in run products
CationFractionsNoTitanite <- filter(CationFractions, Titanite == "N")
CationFractionsTitanite <- filter(CationFractions, Titanite == "Y")
```

### Scatterplot matrices

```{r}
scatterplotMatrix(~ Ti + Ca + InvTK + TK + M, data = CationFractionsTitanite, groups = CationFractionsTitanite$Source, smooth = FALSE)
dev.copy(png,"ScatterPlotMatrixAllExpts.png")
dev.off()   

```

- LEPR compositions show bimodal distribution for melt composition parameter M. 
- LEPR data show different dependence on Flanagan C parameter than our data, so C parameter not included in analysis. 

### Correlation matrix

```{r}
CorrTest <- select(CationFractionsTitanite, TiO2, Pkb, Ca, M:PInvTK)
cor(CorrTest)
```

### Subsets regression
- to find optimal set of parameters 

```{r}
leaps <- regsubsets(TiO2 ~ TK + Pkb + M + FM + InvTK, data = CationFractionsTitanite, nbest = 2) 
plot(leaps, scale = "adjr2")

```

M is in best model fit and FM is not, so use M in model.

### Preliminary model fit

```{r}
RegModel <- CationFractionsTitanite
fit <- lm(TiO2 ~ M + InvTK + Pkb + TK, data = RegModel)
summary(fit)

```

- Drop Pkb, not significant
- TK fits better than InvTK, so drop InvTK

```{r}
fit <- lm(TiO2 ~ M + TK, data = RegModel)
summary(fit)
plot(fit)
```

```{r}
DelExpts <- c("SpDis_1", "SpDis_3", "SpG_28")
RegModel <- filter(RegModel, !(Expt %in%  DelExpts))
fit <- lm(TiO2 ~ M + TK, data = RegModel)
summary(fit)
plot(fit)
```

```{r}
DelExpts <- c("ASI200", "SpDis_2", "DAC")
RegModel <- filter(RegModel, !(Expt %in%  DelExpts))
fit <- lm(TiO2 ~ M + TK, data = RegModel)
summary(fit)
plot(fit)
```


```{r}
DelExpts <- c("HL220-60000", "SH2Na", "SH2K")
RegModel <- filter(RegModel, !(Expt %in%  DelExpts))
fit <- lm(TiO2 ~ M + TK, data = RegModel)
out <- summary(fit)
plot(fit)
```

### Calculate saturation concentrations from model

```{r}
RegModel <- RegModel %>%
  mutate(
   TiO2sat = (out$coefficients[2,1] * M + out$coefficients[3,1] * TK + out$coefficients[1,1]),
    TiO2satSD = TiO2sat * sqrt((out$coefficients[2,2]/ out$coefficients[2,1])^2 + (out$coefficients[3,2] / out$coefficients[3,1])^2),
   lower = TiO2sat - TiO2satSD,
   upper = TiO2sat + TiO2satSD
  )
```

### log-log plot

```{r}
RegModel <- RegModel %>%
  mutate(
    logTiO2 = log10(TiO2),
    logTiO2sat = log10(TiO2sat)
  )
p <- ggplot(RegModel, aes(x = logTiO2, y = logTiO2sat, ymin = log10(lower), ymax = log10(upper), shape = Source, color = Source)) +
  geom_pointrange() +
  geom_abline(slope = 1, intercept = 0) +
  theme(legend.position = "left") +
  ylab(~paste("log"[10], "TiO"[2],"sat")) +
  xlab(~paste("log"[10], "TiO"[2],"meas"))
p1 <- ggMarginal(p, type = "histogram")
print(p1)

```

### Final model fit

Need to delete the two experiments with the lowest TiO2, since they don't fall on the trend and are likely below the detection limit

```{r}
DelExpts <- c("ND-002-18", "ND-002-20")
RegModel <- filter(RegModel, !(Expt %in%  DelExpts))
fit <- lm(TiO2 ~ M + TK, data = RegModel)
out <- summary(fit)
plot(fit)
influencePlot(fit, main="Influence Plot", sub="Circle size is proportional to Cook's distance")

```

Observations depicted as large circles may have a disproportionate influence on the parameter estimates of the model. Points with Hat values above 0.2 have high leverage.

### Test of regression assumptions

```{r}
gvmodel <- gvlma(fit)
summary(gvmodel)
```

"Global Stat" and "Skewness" assumptions not satisfied [@pena2006global]

### Calculate model TiO2 concentrations and errors

```{r}
RegModel <- RegModel %>%
  mutate(
   TiO2sat = (out$coefficients[2,1] * M + out$coefficients[3,1] * TK + out$coefficients[1,1]),
    TiO2satSD = TiO2sat * sqrt((out$coefficients[2,2]/ out$coefficients[2,1])^2 + (out$coefficients[3,2] / out$coefficients[3,1])^2),
   lower = TiO2sat - TiO2satSD,
   upper = TiO2sat + TiO2satSD
  )
```

### Plot model versus measured TiO2

```{r}
fitparams <- c(out$coefficients[1,1], out$coefficients[2,1],  out$coefficients[3,1])
fitparams <- round(fitparams, digits = 4)
p <- ggplot(RegModel, aes(x = TiO2, y = TiO2sat, ymin = lower, ymax = upper, shape = Source, color = Source)) +
  geom_pointrange() +
  geom_abline(slope = 1, intercept = 0) +
  # labs(caption = str_c("TiO2sat = ",fitparams[2], "* M + ", fitparams[3], "* T(K) + ",fitparams[1])) +
  theme(legend.position = "left") +
  ylab(~paste("TiO"[2],"sat")) +
  xlab(~paste("TiO"[2],"meas"))
# add marginal distributions
p1 <- ggMarginal(p, type = "histogram")
print(p1)
ggsave("AllExptsMV2ModelFit.svg", width = 2000, height = 2*618, units = "px", plot = p1)

```

### log-log plot of model fit

```{r}
p <- ggplot(RegModel, aes(x = logTiO2, y = logTiO2sat, ymin = log10(lower), ymax = log10(upper), shape = Source, color = Source)) +
  geom_pointrange() +
  geom_abline(slope = 1, intercept = 0) +
  theme(legend.position = "left") +
  ylab(~paste("log"[10], "TiO"[2],"sat")) +
  xlab(~paste("log"[10], "TiO"[2],"meas"))
p1 <- ggMarginal(p, type = "histogram")
print(p1)
```

### TiO2-M plot

```{r}
ggplot(RegModel, aes(x=M, y=TiO2, color=Source)) +
  geom_point()
```

### TiO2-TK plot

```{r}
ggplot(RegModel, aes(x=TK, y=TiO2, color=Source)) +
  geom_point()
```

### Model output table

```{r}
TiO2AllExptsModelTable <- select(RegModel, Source, Expt, M, TK, TiO2, TiO2sat, TiO2satSD)
write.csv(TiO2AllExptsModelTable, file = "TiO2AllExptsModelTable.csv")
```

### Titanite saturation test

Use Model equation to test whether Expts in "NoTitanite" without rutile are undersaturated. All should have been saturated. 

```{r}
NoTitaniteModel <- filter(CationFractionsNoTitanite, Rutile == "N")
NoTitaniteModel <- select(NoTitaniteModel, Expt, Source, TiO2, TK, M)
NoTitaniteModel$TiO2sat <- predict(fit, NoTitaniteModel)
CompareTiO2 <- select(NoTitaniteModel, Expt, TiO2, TiO2sat)
CompareTiO2 <- CompareTiO2 %>%
  mutate(
    Model_Correct = if_else(TiO2sat > TiO2, "true", "false")
  )
CompareTiO2
CompareTiO2 %>%
  count(Model_Correct)
```

### Save output file and display session info

```{r}
save.image("TitaniteExptsM.RData")
sessionInfo()
```

### References