---
title: 'Titanite solubility: Flanagan + LEPR experiments'
author: "John C. Ayers"
output:
  html_document:
    df_print: paged
    toc: yes
bibliography: "TitaniteReferences.bib"
---

```{r set-global, include=FALSE, echo=FALSE}
# Suppress all error messages
knitr::opts_chunk$set(error = FALSE)
```

```{r setup, include=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, skimr, boot)

```

### Input data

```{r}
AllExpts <- read_csv("AllExptsV3.csv")
# Contains 32 of 39 DF experiments
skimr::skim(AllExpts)
```

### Calculate concentrations of cations

```{r, echo=FALSE}
Cations <- AllExpts %>%
  mutate(
    Si = SiO2/60.0848,
    Al = 2*Al2O3/101.961,
    Na = 2*Na2O/61.979,
    K = 2*K2O/94.194,
    Ca = CaO/56.0794,
    Fe = FeO/71.8464,
    Mg = MgO/40.3044,
    Ti = TiO2/79.8988,
    eCa = (CaO/56.0794 - Al2O3/101.961 + K2O/94.194 + Na2O/61.979),
    CationSum = (Si + Al + Na + K + Ca + Fe + Mg + Ti)
     )
```

### Calculate cation fractions

```{r, echo=FALSE}

CationFractions <- Cations
f <- match("Si", names(CationFractions))
l <- match("Ti", names(CationFractions))
for (i in f:l) {
  CationFractions[(i)] = CationFractions[(i)]/CationFractions$CationSum
}

```

### Calculate melt composition parameters

Formula for calculating melt composition parameter M [@Ryerson1987]

$$M = \frac{(Na + K + 2*Ca)}{(Si * Al)}$$
Formula for calculating melt composition parameter C:

$$C = \frac{10*eCa}{Si * Al}$$ 

where symbols are cation fractions, and excess calcium is defined as

$$ eCa = \frac{(CaO - Al_{2}O_{3} + Na_{2}O + K_{2}O)}{Σ cations}$$

```{r, echo=FALSE}
CationFractions <- CationFractions %>%
  mutate(
    M = (Na + K + 2*Ca)/(Si*Al),
    FM = (Na + K + 2*(Ca + Fe + Mg))/(Si*Al),
    C = (10*eCa)/(Si*Al),
    TK = TC+273.15,
    InvTK = 1/TK,
    PInvTK = (Pkb*1000 - 1)*InvTK,
    lnTiO2 = log(TiO2)
    )
# Filter out experiments with no titanite in run products
CationFractionsNoTitanite <- filter(CationFractions, Titanite == "N")
CationFractionsTitanite <- filter(CationFractions, Titanite == "Y")
```

The methods below are described in section 12.6.2 of [Kabacoff2015]

### Preliminary M model fit

```{r, echo=FALSE}
RegModel <- CationFractionsTitanite
bs <- function(formula, data, indices) {
  d <- data[indices,]
  fit <- lm(formula, data = d)
  return(coef(fit))
}
set.seed(4329)
results <- boot(data = RegModel, statistic = bs, R=1000, formula = lnTiO2 ~ M + InvTK)
print(results)
```

```{r}
rsq <- function(formula, data, indices) {
  d <- data[indices,]
  fit <- lm(formula, data = d)
  return(summary(fit)$r.square)
}
set.seed(4329)
results <- boot(data = RegModel, statistic = rsq, R=1000, formula = lnTiO2 ~ M + InvTK)
print(results) # Mean is only 0.79, about the same as in my preliminary fit using simple lm. 
plot(results) # plots distribution of r-squared values. 
```

### Calculate M model TiO2 concentrations and errors

```{r, echo=FALSE}
# Neither of the following work and I can't figure out why; just get "NA" values
RegModel$lnTiO2sat = results$t0[1] + results$t0[2]*RegModel$M  +results$t0[3]*RegModel$InvTK
RegModel <- RegModel %>%
  mutate(
   lnTiO2sat = results$t0[1] + results$t0[2]*M  +results$t0[3]*InvTK
      )
```

```{r}
ggplot(RegModel, aes(x=lnTiO2, y=lnTiO2sat, shape = Source, color = Source)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme(legend.position = "left") +
  ylab(~paste("TiO"[2],"sat")) +
  xlab(~paste("TiO"[2],"meas"))
```

