---
title: 'Analysis of Titanite solubility experiments'
author: "John C. Ayers"
output:
  html_document:
    df_print: paged
    toc: yes
bibliography: "TitaniteReferences.bib"
---
```{r set-global, include=FALSE}
# Suppress all error messages
knitr::opts_chunk$set(error = FALSE)
```


```{r setup, echo=FALSE, include=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, car, leaps, knitr, skimr, ggExtra)

```

### Import normalized EDS glass analyses

```{r}
TitaniteExpts <- read_csv("TitaniteExptsV3.csv")
skimr::skim(TitaniteExpts)

```

### Convert oxides to cation concentrations

```{r}

GlassCations <- TitaniteExpts %>%
  mutate(
    Si = SiO2 / 60.0848,
    Al = 2 * Al2O3 / 101.961,
    Na = 2 * Na2O / 61.979,
    K = 2 * K2O / 94.194,
    Ca = CaO / 56.0794,
    Fe = FeO / 71.8464,
    Mg = MgO / 40.3044,
    Ti = TiO2 / 79.8988,
    Mn = MnO / 70.9374,
    P = 2 * P2O5 / 141.945,
    eCa = (CaO/56.0794 - Al2O3/101.961 + K2O/94.194 + Na2O/61.979),
    CationSum = (Si + Al + Na + K + Ca + Fe + Mg + Ti + Mn + P)
  )
```

### Calculate cation fractions

```{r}

CationFractions <- select(GlassCations, Expt, Type, SM, TC, Pkb, H2O, TimeH, Titanite, Rutile, TiO2, Si:CationSum)
f <- match("Si", names(CationFractions))
l <- match("P", names(CationFractions))
for (i in f:l) {
  CationFractions[(i)] <- CationFractions[(i)] / CationFractions$CationSum
}
```

### Calculate melt composition parameters

Formula for calculating melt composition parameter M [@Ryerson1987]

$$M = \frac{(Na + K + 2*Ca)}{(Si * Al)}$$

```{r Calculate melt composition parameters}

CationFractions <- CationFractions %>%
  mutate(
    M = (Na + K + 2 * Ca) / (Si * Al),
    FM = (Na + K + 2 * (Ca + Fe + Mg)) / (Si * Al),
    C = (10*eCa)/(Si*Al),
    TK = TC + 273.15,
    InvTK = 1 / TK,
    PInvTK = (Pkb * 1000 - 1) * InvTK
      )
```

### Glass Compositions

```{r}
# Insert table of glass compositions
GlassComps <- select(GlassCations, Expt, TC:TimeH, SiO2:P2O5)
knitr::kable(GlassComps, digits = 2)

```

### Time series plots for titanite-saturated experiments

```{r}

NoTitanite <- filter(CationFractions, Titanite == "N")
CationFractionsTitanite <- filter(CationFractions, Titanite == "Y")
CationFractionsTitanite$Isotherm <- factor(CationFractionsTitanite$TC)
SMplot <- c("GSP + AGV", "PST", "GSP")
for (i in seq_along(SMplot)) {
  Comp <- filter(CationFractionsTitanite, SM == SMplot[i])
  p <- ggplot(Comp, aes(x = TimeH, y = TiO2, shape = Isotherm, color = Isotherm)) +
    geom_point() +
    labs(title = as.name(SMplot[i]))
  print(p)
}

```

Time series a bit concerning, since many isotherms seem to have a negative slope, suggesting that titanite crystallization is kinetically inhibited.

### Correlation matrix

```{r}

CorrTest <- select(CationFractionsTitanite, TiO2, TimeH, Pkb, H2O, Ca, M:PInvTK)
cor(CorrTest)

```

### Scatterplot matrices

```{r}

CationFractionsTitanite$Type <- factor(CationFractionsTitanite$Type)
scatterplotMatrix(~ Ti + Ca + TimeH + TK + M +C, data = CationFractionsTitanite, smooth = FALSE)
```

Glass TiO2 concentration is positively correlated with melt composition parameter M and TK and negatively correlated with time and InvTK.

### Best subsets regression

```{r}

leaps <- regsubsets(TiO2 ~ TK + Pkb + H2O + TimeH + M + FM + C + InvTK, data = CationFractionsTitanite, nbest = 2)
plot(leaps, scale = "adjr2")
# summary(leaps)
```

Since TK and InvTK convey the same information, and theory suggests that titanite solubility should depend on inverse temperature, and TK has a higher correlation coefficient, we include only TK in the analysis. Results of subsets regression suggest that M, TK, Pkb, and TimeH are the significant independent variables.

### Preliminary regression fit

```{r}

fit <- lm(TiO2 ~ M + TK + Pkb + TimeH, data = CationFractionsTitanite)
summary(fit)
```

### Refined regression model

- Pkb and TimeH are not significant, so drop
- Regression of dependent variable TiO2 (wt.%) as a function of independent variables M and TK

```{r}

fit <- lm(TiO2 ~ M + TK, data = CationFractionsTitanite)
summary(fit)
plot(fit)
```

### Final regression model

```{r}
RegModel <- select(CationFractionsTitanite, Expt, Type, TiO2, TK, M)
DelExpts <- c("SpDis_1", "SpDis_15", "SpDis_4")
RegModel <- filter(RegModel, !(Expt %in% DelExpts))
fit <- lm(TiO2 ~ M + TK, data = RegModel)
out <- summary(fit)
out
plot(fit)
```

I think this is our best model. Removing more experiments doesn't significantly improve fit.

### Test of regression assumptions

```{r}
library(gvlma)
gvmodel <- gvlma(fit)
summary(gvmodel)
```

```{r Influence plot and data output}
influencePlot(fit, main = "Influence Plot", sub = "Circle size is proportional to Cook's distance")
write.csv(RegModel, "ExptsInModelV4.csv")
```

Observations depicted as large circles may have a disproportionate influence on the parameter estimates of the model. Points with Hat values above 0.2 have high leverage.

### Model TiO2 concentrations and errors

```{r}

RegModel <- RegModel %>%
  mutate(
    TiO2sat = (out$coefficients[2,1] * M + out$coefficients[3,1] * TK + out$coefficients[1,1]),
    TiO2satSD = TiO2sat * sqrt((out$coefficients[2,2]/ out$coefficients[2,1])^2 + (out$coefficients[3,2] / out$coefficients[3,1])^2),
    lower = TiO2sat - TiO2satSD,
    upper = TiO2sat + TiO2satSD
  )

```

### Plot model versus measured TiO2

```{r}
fitparams <- c(out$coefficients[1,1], out$coefficients[2,1],  out$coefficients[3,1])
fitparams <- round(fitparams, digits = 2)
p <- ggplot(RegModel, aes(x = TiO2, y = TiO2sat, ymin = lower, ymax = upper, shape = Type, color = Type)) +
  geom_pointrange() +
  geom_abline(slope = 1, intercept = 0) +
  labs(caption = str_c("TiO2", " = ",fitparams[2], "* M ", fitparams[3], "* 1/T(K) + ",fitparams[1])) +
  theme(legend.position = "left")
# add marginal distributions
p1 <- ggMarginal(p, type = "histogram")
print(p1)

```

### log-log plot

```{r}
RegModel <- RegModel %>%
  mutate(
    logTiO2 = log10(TiO2),
    logTiO2sat = log10(TiO2sat)
  )
p <- ggplot(RegModel, aes(x = logTiO2, y = logTiO2sat, ymin = log10(lower), ymax = log10(upper), shape = Type, color = Type)) +
  geom_pointrange() +
 # geom_text(label = RegModel$Expt, nudge_x = 0.25, nudge_y = 0.25) +
  geom_abline(slope = 1, intercept = 0) +
  xlim(-1,1.5) +
  ylim(-1,1.5) +
  theme(legend.position = "left") +
  ylab(~paste("log"[10], "TiO"[2][,sat])) +
  xlab(~paste("log"[10], "TiO"[2][,meas]))
p1 <- ggMarginal(p, type = "histogram")
print(p1)

```

### TiO2 vs. M plot

```{r}

ggplot(RegModel, aes(x = M, y = TiO2, shape = Type, color = Type)) +
  geom_point()
```

### TiO2-TK plot

```{r}
ggplot(RegModel, aes(x = TK, y = TiO2, color = Type)) +
  geom_point()
```

### TiO2-TK plot

```{r}

ggplot(RegModel, aes(x = TK, y = TiO2, color = Type)) +
  geom_point()
```

### Model output table

```{r}
TiO2ModelTable <- select(RegModel, Expt, M, TK, TiO2, TiO2sat,TiO2satSD)
kable(TiO2ModelTable, digits = 2, format = "html",
             col.names = c("Expt.", "M", "T(K)", "TiO~2~", "TiO~2,sat~", "1$\\sigma$"))
write.csv(TiO2ModelTable, file = "TiO2ModelTable.csv")
```

### Undersaturation test

Use Model equation to test whether Expts in "NoTitanite" are undersaturated

```{r}
NoTitaniteModel <- filter(NoTitanite, Rutile == "N")
NoTitaniteModel <- select(NoTitaniteModel, Expt, Type, TiO2, TK, M)
NoTitaniteModel$TiO2sat <- predict(fit, NoTitaniteModel)
CompareTiO2 <- select(NoTitaniteModel, Expt, TiO2, TiO2sat)
CompareTiO2 <- CompareTiO2 %>%
  mutate(
    Model_Correct = if_else(TiO2sat > TiO2, "true", "false")
  )
CompareTiO2
CompareTiO2 %>%
  count(Model_Correct)

```
### See how model equation fits LEPR data

### Convert oxides to cation concentrations

```{r}
LEPR_Expts <- read_csv("LEPR_Titanite_ExptsV2.csv")
LEPR_Cations <- LEPR_Expts %>%
  mutate(
    Si = SiO2 / 60.0848,
    Al = 2 * Al2O3 / 101.961,
    Na = 2 * Na2O / 61.979,
    K = 2 * K2O / 94.194,
    Ca = CaO / 56.0794,
    Fe = FeO / 71.8464,
    Mg = MgO / 40.3044,
    Ti = TiO2 / 79.8988,
    CationSum = (Si + Al + Na + K + Ca + Fe + Mg + Ti)
  )
```

### Calculate cation fractions

```{r}

LEPRCationFractions <- select(LEPR_Cations, Expt, TC, TiO2, Si:CationSum)
f <- match("Si", names(LEPRCationFractions))
l <- match("Ti", names(LEPRCationFractions))
for (i in f:l) {
  LEPRCationFractions[(i)] <- LEPRCationFractions[(i)] / LEPRCationFractions$CationSum
}
```


```{r}

LEPRCationFractions <- LEPRCationFractions %>%
  mutate(
    M = (Na + K + 2 * Ca) / (Si * Al),
    TK = TC + 273.15,
          )
```

```{r}
RegModel <- select(LEPRCationFractions, Expt, TiO2, TK, M)
RegModel <- RegModel %>%
  mutate(
    TiO2sat = (out$coefficients[2,1] * M + out$coefficients[3,1] * TK + out$coefficients[1,1]),
    TiO2satSD = TiO2sat * sqrt((out$coefficients[2,2]/ out$coefficients[2,1])^2 + (out$coefficients[3,2] / out$coefficients[3,1])^2),
    lower = TiO2sat - TiO2satSD,
    upper = TiO2sat + TiO2satSD
  )
```

### Plot model versus measured TiO2

```{r}
fitparams <- c(out$coefficients[1,1], out$coefficients[2,1],  out$coefficients[3,1])
fitparams <- round(fitparams, digits = 2)
p <- ggplot(RegModel, aes(x = TiO2, y = TiO2sat, ymin = lower, ymax = upper)) +
  geom_pointrange() +
  geom_abline(slope = 1, intercept = 0) 
# add marginal distributions
# p1 <- ggMarginal(p, type = "histogram")
print(p)

```

### Statistics

```{r}
fit <- lm(TiO2sat ~ TiO2, data = RegModel)
summary(fit)
```

### Model output table

```{r}
TiO2ModelTable <- select(RegModel, Expt, M, TK, TiO2, TiO2sat,TiO2satSD)
kable(TiO2ModelTable, digits = 2, format = "html",
             col.names = c("Expt.", "M", "T(K)", "TiO~2~", "TiO~2,sat~", "1$\\sigma$"))
```

### Save output file and display session info

```{r}
save.image("TitaniteExptsM.RData")
sessionInfo()
```

# References
