---
title: "Titanite solubility experiments: Analysis of EDS data"
output: html_notebook
---

```{r}
library(tidyverse, warn.conflicts = FALSE)
# Import normalized EDS glass analyses
TitaniteExpts <- read_csv("TitaniteExpts.csv")
```
```{r}
TitaniteCations <- TitaniteExpts %>%
  mutate(
    Si = SiO2/60.0848,
    Al = 2*Al2O3/101.961,
    Na = 2*Na2O/61.979,
    K = 2*K2O/94.194,
    Ca = CaO/56.0794,
    Fe = FeO/71.8464,
    Mg = MgO/40.3044,
    Ti = TiO2/79.8988,
    Mn = MnO/70.9374,
    P = 2*P2O5/141.945,
    eCa = (CaO/56.0794 - Al2O3/101.961 + K2O/94.194 + Na2O/61.979),
    CationSum = (Si + Al + Na + K + Ca + Fe + Mg + Ti + Mn + P)
             )
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.
```{r}
# Calculate cation fractions
CationFractions <- select(TitaniteCations, Expt, TC, Si:eCa, CationSum)
f <- match("Si", names(CationFractions))
l <- match("eCa", names(CationFractions))
for (i in f:l) {
  CationFractions[(i)] = CationFractions[(i)]/CationFractions$CationSum
}

```


```{r}
CationFractions <- CationFractions %>%
  mutate(
    M = (Na + K + 2*Ca)/(Si*Al),
    FM = (Na + K + 2*(Ca + Fe + Mg))/(Si*Al),
    C = (10*eCa)/(Si*Al),
    InvTK = 1/(TC+273.15)
    )
```
When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
