---
title: 'Titanite solubility: Flanagan + LEPR experiments'
author: "John C. Ayers"
output:
  html_document:
    df_print: paged
    toc: yes
bibliography: "TitaniteReferences.bib"
---

This program imports the file AllExptsV3.csv containing silicate glass compositions from:
1. titanite growth experiments conducted at Vanderbilt University by graduate student Danny Flanagan.
2. titanite dissolution experiments conducted at RPI by Rick Ryerson and Bruce Watson. 
3. phase equilibrium experiments from the LEPR database [@Hirschmann2008] that produced titanite.
Thermodynamics suggests that a titanite solubility equation should have the form:

ln (TiO2)melt = -ΔH°r/RT + ΔS°r/R - ΔVr(P-1)/RT – ln γTiO2melt – k1M

ln(TiO2)melt, the natural log of the concentration of TiO2 required for titanite saturation in the melt, is the dependent variable. The independent variables are 1/T, -ΔV(P-1)/T and melt composition M. Subsets regression is used to select the independent variables that produce the best model. Iterative rejection of outliers leads to a refined multiple linear regression model. The model is validated by importing data for titanite-undersaturated melts from the LEPR database (files LEPR_ExperimentsV2.csv and LEPR_LiquidAnhyV4.csv) and calculating the percent of experiments that are correctly predicted to be undersaturated in titanite. 


```{r set-global, include=FALSE}
# Suppress all error messages
knitr::opts_chunk$set(error = FALSE)
```

```{r setup, include=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, car, leaps, knitr, skimr, ggExtra, gvlma, svglite)

```

### Input data

```{r}
AllExpts <- read_csv("AllExptsV3.csv")
# Contains 32 of 39 DF experiments
skimr::skim(AllExpts)
```

### Calculate concentrations of cations

```{r}
Cations <- AllExpts %>%
  mutate(
    Si = SiO2/60.0848,
    Al = 2*Al2O3/101.961,
    Na = 2*Na2O/61.979,
    K = 2*K2O/94.194,
    Ca = CaO/56.0794,
    Fe = FeO/71.8464,
    Mg = MgO/40.3044,
    Ti = TiO2/79.8988,
    eCa = (CaO/56.0794 - Al2O3/101.961 + K2O/94.194 + Na2O/61.979),
    CationSum = (Si + Al + Na + K + Ca + Fe + Mg + Ti),
    lnTiO2 = log(TiO2),
    Al2O3m = Al2O3/101.961,
    Na2Om = Na2O/61.979,
    K2Om = K2O/94.194,
    CaOm = CaO/56.0794
     )
```

### Calculate cation fractions

```{r}

CationFractions <- Cations
f <- match("Si", names(CationFractions))
l <- match("Ti", names(CationFractions))
for (i in f:l) {
  CationFractions[(i)] = CationFractions[(i)]/CationFractions$CationSum
}

```

### Calculate melt composition parameters

Formula for calculating melt composition parameter M [@Ryerson1987]

$$M = \frac{(Na + K + 2*Ca)}{(Si * Al)}$$

```{r}
CationFractions <- CationFractions %>%
  mutate(
    M = (Na + K + 2*Ca)/(Si*Al),
    FM = (Na + K + 2*(Ca + Fe + Mg))/(Si*Al),
    C = (10*eCa)/(Si*Al),
    A_CNK = Al2O3m/(CaOm + Na2Om + K2Om),
    A_NK = Al2O3m/(Na2Om + K2Om),
    A_C = Al2O3m/CaOm,
    TK = TC+273.15,
    InvTK = 1/TK,
    PInvTK = (Pkb*1000 - 1)*InvTK
    )
# Filter out experiments with no titanite in run products glasses
CationFractionsNoTitanite <- filter(CationFractions, Titanite == "N")
CationFractionsTitanite <- filter(CationFractions, Titanite == "Y")
```

### Correlation matrix

```{r}
CorrTest <- select(CationFractionsTitanite, TiO2, lnTiO2, Pkb, Ca, M:PInvTK, A_NK, A_CNK, A_C)
cor(CorrTest)
```

-   FM parameter does not correlate as well as M or C, so drop FM.
-   Pkb and PInvTK show poor correlations, so drop from model.

### Subsets regression

-   to find optimal set of parameters

```{r}
leaps <- regsubsets(TiO2 ~ TK + M + C + InvTK, data = CationFractionsTitanite, nbest = 2) 
plot(leaps, scale = "adjr2")

```

-   M is in best model fit and C is not, so use M and not C.

### Scatterplot matrices

```{r}
scatterplotMatrix(~ Ti + A_CNK + InvTK + TK + C + M, data = CationFractionsTitanite, groups = CationFractionsTitanite$Source, smooth = FALSE)
```

-   LEPR compositions show bimodal distribution for melt composition parameters M and C.

### Preliminary M model fit

```{r}
RegModel <- CationFractionsTitanite
fit <- lm(TiO2 ~ M + InvTK + TK, data = RegModel)
summary(fit)

```

-   TK fits better than InvTK, so drop InvTK

### Refined M model fit

```{r, results='hide', fig.show='hide'}
fit <- lm(TiO2 ~ M + TK, data = RegModel)
summary(fit)
plot(fit)
```

```{r, results='hide'}
DelExpts <- c("SpDis_1", "SpDis_3", "SpG_28")
RegModel <- filter(RegModel, !(Expt %in%  DelExpts))
fit <- lm(TiO2 ~ M + TK, data = RegModel)
summary(fit)
```

```{r, results='hide'}
DelExpts <- c("ASI200", "SpDis_2", "DAC")
RegModel <- filter(RegModel, !(Expt %in%  DelExpts))
fit <- lm(TiO2 ~ M + TK, data = RegModel)
summary(fit)
```

```{r, results='hide'}
DelExpts <- c("HL220-60000", "SH2Na", "SH2K")
RegModel <- filter(RegModel, !(Expt %in%  DelExpts))
fit <- lm(TiO2 ~ M + TK, data = RegModel)
out <- summary(fit)
out
```

```{r}
Mcoef <- out$coefficients[2,1]
Tcoef <- out$coefficients[3,1]
Int <- out$coefficients[1,1]
RegModel <- RegModel %>%
  mutate(
   TiO2sat = (Mcoef * M + Tcoef * TK + Int)
      )
```

### preliminary M log-log plot

```{r}
p <- ggplot(RegModel, aes(x = log10(TiO2), y = log10(TiO2sat), shape = Source, color = Source)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme(legend.position = "left") +
  ylab(~paste("log"[10], "TiO"[2],"sat")) +
  xlab(~paste("log"[10], "TiO"[2],"meas"))

```

### Final M model fit

Need to delete the two experiments with the lowest TiO2, since they don't fall on the trend and are likely below the detection limit

```{r, results='hide'}
DelExpts <- c("ND-002-18", "ND-002-20")
RegModel <- filter(RegModel, !(Expt %in%  DelExpts))
fit <- lm(TiO2 ~ M + TK, data = RegModel)
out <- summary(fit)
plot(fit)
influencePlot(fit, main="Influence Plot", sub="Circle size is proportional to Cook's distance")

```

Observations depicted as large circles may have a disproportionate influence on the parameter estimates of the model. Points with Hat values above 0.2 have high leverage.

### Test of M model regression assumptions

```{r}
gvmodel <- gvlma(fit)
summary(gvmodel)
```

All assumptions satisfied [@pena2006global]

### Calculate M model TiO2 concentrations and errors

```{r}
# Model output table provides standard errors; convert to standard deviation by multiplying by sqrt of n = # of samples
n <- nrow(RegModel)
Mcoef <- out$coefficients[2,1]
Mcoefsd <- out$coefficients[2,2]*sqrt(n)
Tcoef <- out$coefficients[3,1]
Tcoefsd <- out$coefficients[3,2]*sqrt(n)
Int <- out$coefficients[1,1]
RegModel <- RegModel %>%
  mutate(
   TiO2sat = (Mcoef * M + Tcoef * TK + Int),
    TiO2satSD = sqrt(Mcoef^2*Mcoefsd^2 + Tcoef^2*Tcoefsd^2),
   lower = TiO2sat - TiO2satSD,
   upper = TiO2sat + TiO2satSD,
    )
```

### Plot M model TiO2 saturation concentration versus measured TiO2

```{r}
p <- ggplot(data = RegModel, aes(x = TiO2, y = TiO2sat, ymin = lower, ymax = upper, shape = Source, color = Source)) +
  geom_pointrange() +
  geom_abline(slope = 1, intercept = 0) +
  theme(legend.position = "left") +
  ylab(~paste("TiO"[2],"sat")) +
  xlab(~paste("TiO"[2],"meas"))
# add marginal distributions
p1 <- ggMarginal(p, type = "histogram")
print(p1)
ggsave("AllExptsM_ModelFit.svg", width = 2000, height = 2*618, units = "px", plot = p1)

```

### final log-log plot of M model fit

```{r}
p <- ggplot(RegModel, aes(x = log10(TiO2), y = log10(TiO2sat), ymin = log10(lower), ymax = log10(upper), shape = Source, color = Source)) +
  geom_pointrange() +
  geom_abline(slope = 1, intercept = 0) +
  theme(legend.position = "left") +
  ylab(~paste("log"[10], "TiO"[2],"sat")) +
  xlab(~paste("log"[10], "TiO"[2],"meas"))
p1 <- ggMarginal(p, type = "histogram")
print(p1)
```

### TiO2-M plot

```{r}
RegModel %>%
ggplot(aes(x=M, y=TiO2, fill = TC)) +
  geom_point(size = 4, shape = 22, color = "black") +
   theme_bw() +
  ylab(~paste("TiO"[2],"sat")) +
 scale_fill_gradientn(colours = c("darkred", "orange", "yellow", "white"))
  
```

### TiO2-TK plot

```{r}
ggplot(RegModel, aes(x=TK, y=TiO2, color=M)) +
  geom_point()
```

### TiO2 - A/CNK plot

```{r}
ggplot(RegModel, aes(x=A_CNK, y=TiO2, color=TK)) +
  geom_point()
```

### Model output table

```{r}
TiO2AllExptsModelTable <- select(RegModel, Source, Expt, TiO2, CaO, M, FM, A_CNK, A_C, TK, TiO2sat, TiO2satSD)
write.csv(TiO2AllExptsModelTable, file = "TiO2AllExptsM_ModelTable.csv")
DFGlassComps <- filter(RegModel, Source != "LEPR") %>%
  select(Expt, SiO2:TiO2, M, FM, A_CNK, A_C)
write.csv(DFGlassComps, file = "DFGlassCompositions.csv")

```

### M model titanite saturation test

Use M model equation to test whether LEPR expts without titanite or rutile are undersaturated.

```{r}
LEPR_Expts <- read_csv("LEPR_ExperimentsV2.csv")
LEPRLiqs <- read_csv("LEPR_LiquidAnhyV4.csv")
LEPR <- inner_join(LEPR_Expts, LEPRLiqs, by = "Index")
LEPRnoTitRut <- filter(LEPR, !(grepl("tit|rut", Phases))) %>%
  mutate(
    FeOt = FeO + (2*71.85*Fe2O3)/159.69,
    TK = T_C + 273.15,
    Tot = SiO2 + TiO2 + Al2O3 + FeOt + MgO + CaO + Na2O + K2O) %>%
  filter(Tot > 90)
LEPRnoTitRut <- select(LEPRnoTitRut, Experiment.x, TK, SiO2, TiO2, Al2O3, FeOt, MgO, CaO, Na2O, K2O, Tot)   %>%
  mutate(
    SiO2 = 100*SiO2/Tot,
    TiO2 = 100*TiO2/Tot,
    Al2O3 = 100*Al2O3/Tot,
    FeOt = 100*FeOt/Tot,
    MgO = 100*MgO/Tot,
    CaO = 100*CaO/Tot,
    Na2O = 100*Na2O/Tot,
    K2O = 100*K2O/Tot,
    Tot = SiO2 + TiO2 + Al2O3 + FeOt + MgO + CaO + Na2O + K2O
  )
```

#### Calculate concentrations of cations

```{r}
LEPRnoTitRutCations <- LEPRnoTitRut %>%
  mutate(
    Si = SiO2/60.0848,
    Al = 2*Al2O3/101.961,
    Na = 2*Na2O/61.979,
    K = 2*K2O/94.194,
    Ca = CaO/56.0794,
    Fe = FeOt/71.8464,
    Mg = MgO/40.3044,
    Ti = TiO2/79.8988,
    eCa = (CaO/56.0794 - Al2O3/101.961 + K2O/94.194 + Na2O/61.979),
    CationSum = (Si + Al + Na + K + Ca + Fe + Mg + Ti)
     )
```

#### Calculate cation fractions

```{r}

LEPRnoTitRutCationFractions <- LEPRnoTitRutCations
f <- match("Si", names(LEPRnoTitRutCationFractions))
l <- match("Ti", names(LEPRnoTitRutCationFractions))
for (i in f:l) {
  LEPRnoTitRutCationFractions[(i)] = LEPRnoTitRutCationFractions[(i)]/LEPRnoTitRutCationFractions$CationSum
}

```

#### Calculate melt composition parameters

```{r}
LEPRnoTitRutCationFractions <- LEPRnoTitRutCationFractions %>%
  mutate(
    M = (Na + K + 2*Ca)/(Si*Al),
    C = (10*eCa)/(Si*Al),
    )
```

#### Compare model and observations

```{r}
NoTitaniteModelM <- select(LEPRnoTitRutCationFractions, Experiment.x, TiO2, TK, M)
NoTitaniteModelM$TiO2sat <- predict(fit, NoTitaniteModelM)
CompareTiO2 <- select(NoTitaniteModelM, Experiment.x,  TiO2, TiO2sat)
CompareTiO2 <- CompareTiO2 %>%
  mutate(
    Model_Correct = if_else(TiO2sat > TiO2, "true", "false")
  )
CompareTiO2
CompareTiO2 %>%
  count(Model_Correct)
```

-   M model correctly predicts undersaturation in 95% of experiments.

### Save output file and display session info

```{r}
save.image("TitaniteAllExptsM.RData")
sessionInfo()
```

### References
